---
title: "Tissue-specific CCR8 distribution"
author: "Hyun Yong Jin"
date: "August 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Tissue specific CCR8 expression distribution profile from GSE108989(CRC) and GSE99254
##Start from merged files from previous GSE108989 and GSE99254 analysis.
##Each file contains only specific population (#CD4_C12-CCR8 cluster from GSE108989 or #CD4_C9-CTLA4 from GSE99254) of exhausted CD4 T cell population from each dataset.


## dintinc of tidyverse to show the unique list in the column
## violin plot
## cowplot


```{r}
library(tidyverse)

crc<- read.csv("merged_ccr8_GSE108989_CRC.csv", header=T) #CD4_C12-CCR8 cluster from GSE108989
nsclc<- read.csv("merged_ctla4_GSE99254_NSCLC.csv", header=T) #CD4_C9-CTLA4 from GSE99254

#Sanity test
dim(crc)
dim(nsclc)

crc[c(1:5), c(1:12)]
nsclc[c(1:5), c(1:12)]


#Analysis of CRC

distinct(crc, SampleType) #Show unique values (all clusters)

#PTC(CD8+ T cells from peripheral blood), NTC(CD8+T cells from adjacent normal colonrectal tissues) ,TTC (CD8+ T cells from tumor), 
#PTH(CD3+, CD4+ and CD25- T cells from peripheral blood),NTH(CD3+, CD4+ and CD25- T cells from adjacent normal colonrectal tissues),TTH(CD3+, CD4+ and CD25- T cells from tumor), 
#PTR(CD3+, CD4+ and CD25high T cells from peripheral blood), NTR(CD3+, CD4+ and CD25high Tcells from adjacent normal colonrectal tissues), TTR(CD3+, CD4+ and CD25high T cells from tumor), 
#PTY(CD3+, CD4+ and CD25mediate T cells from peripheral blood), NTY(CD3+, CD4+ and CD25mediate T cells from adjacent normal colonrectal tissues), TTY(CD3+, CD4+ and CD25medate T cells from tumor), 
#PP7(CD3+, CD4+ T cells from peripheral blood), NP7(CD3+, CD4+ T cells from adjacent normal colonrectal tissues), TP7(CD3+, CD4+ T cells from tumor).
#I will focus on PTR, NTR and TTR population. 

crc.p <- crc %>%
  select(SampleType, CCR8) %>%
  filter(SampleType == "PTR") #2 cells

crc.n <- crc %>%
  select(SampleType, CCR8) %>%
  filter(SampleType == "NTR") #42 cells

crc.t <- crc %>%
  select(SampleType, CCR8) %>%
  filter(SampleType == "TTR") #845 cells





#Analysis of NSCLC

distinct(nsclc, tissueType) #Show unique values (all clusters)

nsclc.p <- nsclc %>%
  select(tissueType, CCR8) %>%
  filter(tissueType == "Peripheral blood (P)") #12 cells

nsclc.n <- nsclc %>%
  select(tissueType, CCR8) %>%
  filter(tissueType == "Normal (N)") #59 cells

nsclc.t <- nsclc %>%
  select(tissueType, CCR8) %>%
  filter(tissueType == "Tumor-infiltrating (T)") #868 cells

```



## Generating graph
### Graph1: cell number incidence of of n, p and t population
### Graph2: violin plot of CCR8, each populations. 



```{r}

library(ggplot2)
library(scales) # to format the labels in percent
library(ggrepel)
library(cowplot)

# Graph1: cell number incidence of of n, p and t population

df.crc <-data.frame(
  group = c("Peripheral", "Normal", "Tumor"),
  value = c(nrow(crc.p), nrow(crc.n), nrow(crc.t))
)

pie.crc <- ggplot(df.crc, aes(x="", y=value, fill=group))+
  geom_bar(width =1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("Tissue Origin of \n CD4_C12-CCR8 cluster in CRC") +
  xlab("# of Cells") + ylab("")

plot(pie.crc)
#Save as 6x6

df.nsclc <-data.frame(
  group = c("Peripheral", "Normal", "Tumor"),
  value = c(nrow(nsclc.p), nrow(nsclc.n), nrow(nsclc.t))
)

pie.nsclc <- ggplot(df.nsclc, aes(x="", y=value, fill=group))+
  geom_bar(width =1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("Tissue Origin of \n CD4_C9-CTLA4 in NSCLC") +
  xlab("# of Cells") + ylab("")

plot(pie.nsclc)


#Graph2: violin plot of CCR8, each populations. This will show cut-off values of CCR8 positive and negative.
#Ideally I want to generate CCR8 distribution from three different tissue in one figure using single script. 


# Filter multiple: https://stackoverflow.com/questions/25647470/filter-multiple-values-on-a-string-column-in-dplyr

categories <- c("PTR", "NTR", "TTR")
crc.plot <- crc %>%
  select(SampleType, CCR8) %>%
  filter(SampleType %in% categories) #1042 obs to 889 obs

head(crc.plot)

v.crc <- ggplot(crc.plot, aes(x=SampleType, y=CCR8, color=SampleType)) + 
  geom_violin(trim=FALSE)+
  geom_jitter(shape=16, position=position_jitter(0.2))+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCR8 expression of \n CD4_C12-CCR8 cluster in CRC")


plot(v.crc)
#Save as 6x4



categories2 <- c("Peripheral blood (P)", "Normal (N)", "Tumor-infiltrating (T)")
nsclc.plot <- nsclc %>%
  select(tissueType, CCR8) %>%
  filter(tissueType %in% categories2) #939 obs to 939 obs. No other tissue types were found

head(nsclc.plot)

v.nsclc <- ggplot(nsclc.plot, aes(x=tissueType, y=CCR8, color=tissueType)) + 
  geom_violin(trim=FALSE)+
  geom_jitter(shape=16, position=position_jitter(0.2))+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCR8 expression of \n CD4_C9-CTLA4 in NSCLC")


plot(v.nsclc)




```


## cowplot to generate final figure

```{r cowplot}
#cowplot to show all results side-by-side
#https://stackoverflow.com/questions/1249548/side-by-side-plots-with-ggplot2

library(cowplot)

p<-plot_grid(pie.crc, pie.nsclc, v.crc, v.nsclc, labels = "AUTO")
ggsave("all_plots_combined.pdf", p)

plot(p)


```


