---
title: "Zemin_NSCLC_GSE99254-CCR8_Analaysis"
author: "Hyun Yong Jin"
date: "August 16, 2019"
output: html_document
---

Analysis was done as same as GSE108989 ver1.1 (Zemin/s CRC)
Ver1.1 as of 20190808 
  Code readability has been improved.
Ver1.2 as of 20190816
  distinct(merged, majorCluster) #This will show list of cluster in the table. 


Downloaded raw table from GSE99254. This table contained normalized gene expression (12415 genes) of single cells (11771 cells).
Downloaded table from Tamatoa. This table includes individual cell ID and cluster information, has more cells (12346). However, unlike CRC dataset, the cluster contains many non-categorized cells (called "nan").
Merged two tables. Now I have individual cell (11769) with gene expression profile (12414), removed one un-assigned gene).
Select cells assigned to CD4_C8-FOXP3. Down to 427 cells.
Starting from the CCR8 cluster, I separated the individual cells into two groups. Cells belong to CCR8hi (log2 >5, 52 cells) and CCR8 low (1<log2<3, 11 cells). A
From this point on, I treated the individual cells from CCR8hi group (52 cells) and low group (11 cells) as biological replicates for calculating statistics.
I calculated mean, FC, SD, p values and other statistics per individual genes.
From this stat(stat_all.csv), I selected FC > 5 and p values <0.01 genes. This table is attached, showing upregulated gene list in CCR8hi cells. 
From this stat(stat_all.csv), I selected FC<0.2 genes and p values <0.2. 




```{r global_options}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Output/',
                     warning=FALSE)
```

##Download Data

```{r download data}
#GSE99254
library(GEOquery)

getGEOSuppFiles('GSE99254', fetch_files = FALSE) #Check to see what is in the supplement files

#There are three normalization data. Based on their method section, norm.centered.txt.gz is the most relevant dataset. 
#Download all, and select "1 GSE108989_CRC.TCell.S10805.norm.centered.txt.gz"

getGEOSuppFiles('GSE99254') #All three files were downloaded in sub-folder /GSE99254

```

## Analysis from the downloaded table

```{r Merging row data with Tamatoa data}
## Row data contains gene ID and expression level of each genes. However, cluster information and selection of cells (excluding low quality cells) information is missing. In contrast, downloaded file from Tamatoa contains cluster info, and lower number of cells, presumably removed the low quality cells.

## Idea: Both table contains cell-ID as common identifier. Merge the two table together for downstream analysis. 

library(tidyverse)

tab <- read.delim("GSE99254/GSE99254_NSCLC.TCell.S11769.norm.centered.txt.gz") #Load file as tab delimited txt.

dim(tab)
tab[c(1:10), c(1:20)]


#Start testing with small scale example. 
tab_test<- tab[c(1:10), c(1:5)]
n1 <- tab_test$geneSymbol
t.tab_test <- as.data.frame(t(tab_test[,-c(1:2)])) #remove 1st and 2nd column and transpose
colnames(t.tab_test) <- n1
str(t.tab_test)


#Transpose the original tab table.
n2 <- tab$geneSymbol
t.tab <- as.data.frame(t(tab[,-c(1:2)])) #remove 1st and 2nd column and transpose
colnames(t.tab) <- n2
t.tab[c(1:5), c(1:6)] #sanity test


#Next is to combine t.tab with data from tamatoa
id<-read.csv("Tamatoa/identifier-cluster_matching.csv", header=T)

#Now issue is the identifier in t.tab is rowname (without header) and id is in column. Both table is data.frame format. 
#to merge tables, the cell id in t.tab has to be assigned.

rownames_test <- tibble::rownames_to_column(t.tab_test,"VALUE") #test with small table
t.tab <-tibble::rownames_to_column(t.tab, "cell_names") #t.tab was overwritten but column was assigne. 

test <- !is.na(names(t.tab)) #64th column nas NA (not assigned) header. Only one missing header. All other headers were fine. 

#Let's remove column with header NA from t.tab
t.tab <- t.tab[test] #re-assign only TRUE values

dim(t.tab) #10805 x 12547 (12548 before removing NA)
dim(id) #7172 x 9

#Inner_join will merge table based on common identifier in the same column name. This function is part of dplyr
t.tab$cell_names[1:10]
class(t.tab$cell_names) #character
id$cell_names[1:10]
class(id$cell_names) #factor

#id$cell_names should be converted to character. For example, second column tSNE1 has numeric value.
#https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters

i <- sapply(id, is.factor)
id[i] <- lapply(id[i], as.character)
class(id$cell_names) #Now the cell_names column is converted to character


#Merging step.
#There are a few issues so I resolved them.
t.tab$cell_names[10] #Id was connected by dot
id$cell_names[10] #Id was connected by hyphen

#convert hyphen to dot in cell_names columne in id
?gsub #pattern matching and replacement
id$cell_names <- gsub("-", ".", id$cell_names)

#Merge two table and excluded cells from no common id. Used inner_join
merged <- inner_join(id, t.tab, by = "cell_names") 
dim(merged) #7172 12555
#Note that id file (from tamatoa) has 7172, and row file has 10805 cells. Among them, 7172 was overlapped. I assume the 3000 cells were removed due to the low expression. 


merged[c(1:8), c(1:12)] #Sanity test.
write.csv(merged, file="merged_all.csv")

#Selection of CD4_C8_FOXP3 cluster only
distinct(merged, majorCluster) #Show unique values (all clusters)
merged_foxp3only<- merged %>% filter(majorCluster =="CD4_C8-FOXP3")
dim(merged_foxp3only) # 427 cells

write.csv(merged_foxp3only, file="merged_foxp3.csv")
#Export the merged dataset.



```



```{r Cluster Foxp3 only}

#How about rowVars? But rowVars detects the most variable genes between individual replicates.

merged_foxp3only[c(1:50), c(1:12)] 
tbl <- merged_foxp3only %>%
  select(-cell_names, -tSNE1, -tSNE2, -majorCluster, -patient, -tissueType, -cellType, -cyclePhase, -Units) #select the necessary columnes only

tbl[c(1:20), c(1:10)]

#transpose data and maintain the first column as header
t_tbl <- as.data.frame(t(tbl)) #Transposing number only is much faster.
colnames(t_tbl) <- merged_foxp3only$cell_names #Add colname back

t_tbl[c(1:20), c(1:10)]
dim(t_tbl)

#Note that this filter is based on log2 value.
ccrhi <- tbl %>% filter(CCR8>5) %>% arrange(desc(CCR8)) #52 obs
ccrlo <- tbl %>% filter(CCR8>1 & CCR8 <3) %>% arrange(desc(CCR8)) #11 obs

#Now treat individual cells as individual replicate.
#Transpose the table
tccrhi <- as.data.frame(t(ccrhi))
tccrlo <- as.data.frame(t(ccrlo))

#Also note that stat should be done in linear values.
#Writing quick function to make the whole table to linear values
lin <- function (x, na.rm=FALSE) (2^x)
lin_tccrhi <- lin(tccrhi)
lin_tccrlo <- lin(tccrlo)

#Make sure both table contains same numbers of genes. 12546
dim(lin_tccrhi)
dim(lin_tccrlo)

#Calculate fold change CCRhi/CCRlo
datFC <- t_tbl %>%
  rownames_to_column("gene_name") %>%
  mutate(hi_mean = rowMeans(lin_tccrhi)) %>%
  mutate(lo_mean = rowMeans(lin_tccrlo)) %>%
  mutate(FC=hi_mean/lo_mean)

datFC[c(1:5), c(1:10)] #first column becomes gene_name

stat <- datFC %>% 
  select(gene_name, hi_mean, lo_mean, FC) #extracting FC stats only. with rownames

#Calculate SD
library(genefilter)
datFCSD <- stat %>%
  mutate(hi_SD = rowSds(lin_tccrhi)) %>%
  mutate(lo_SD = rowSds(lin_tccrlo))

#Writing for-loop to calculate t-test in row-wise.
#See my 20190110 IBD analysis as reference

library(broom) #for tidy function
testresults <- vector("list", nrow(datFCSD))

#Start for-loop. Takes some time.
for (j in seq(nrow(datFCSD))) {
  testresults[[j]] <-tidy(t.test(as.data.frame(lin_tccrhi[j,]), as.data.frame(lin_tccrlo[j,])))
}

t_stats = do.call(rbind, testresults)
head(t_stats)

all_stats <-bind_cols(datFCSD, t_stats)

dim(all_stats)
head(all_stats) 
#Note that my manucal mean calculation and calculation of tidy (estimate1& estimate2) is identical. 


write.csv(all_stats, file="GSE99254_stat_all.csv") #Summarized stat table. 

#Select genes with high fold change and significant p values
up_in_ccr8hi <- all_stats %>%
  select(-estimate1, -estimate2) %>%
  filter(p.value <0.01 & FC >5) %>%
  arrange (desc(FC))
  
down_in_ccr8hi <- all_stats %>%
  select(-estimate1, -estimate2) %>%
  filter(p.value <0.2 & FC <0.2) %>%
  arrange (FC) #default is ascending order

write.csv(up_in_ccr8hi, file="GSE99254_5-fold_up_in_ccr8hi_sig.csv") #Summarized stat table.
write.csv(down_in_ccr8hi, file="GSE99254_5-fold_down_in_ccr8hi.csv") #Summarized stat table. 

#Comments: Many cells with low abundent TPM contains exactly same values. Statistics from these values may not represent true statistics. Removing cells with low abundant values are not feasible because essentially all the cells contain low abundant mRNAs. In these case, fold-change could be more reliable values.  


```


## Visualization


```{r ggplot2}
#Tried to find our Ruggero lab volcano plot + ggrepel script I wrote. 
#I would generate a seperate Markdown file for volcano/ggrepel combination

```


