---
title: "Tissue-specific CCR8 expression and Tissue distribution of select CD4 clusters"
author: "Hyun Yong Jin"
date: "August 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tissue specific CCR8 expression distribution from GSE108989(CRC) and GSE99254 (NSCLC). May investigate other genes of interests including Cxorf21.

Start from merged_all files from previous GSE108989 and GSE99254 analysis.

Select a subset of CD4 population, focusing on #CD4_C12-CCR8 cluster from GSE108989 or #CD4_C9-CTLA4 from GSE99254.Other Treg population and naive T cells (CCR7+ CD4) was also selected as control for tumor Treg cells. 

"dintinc" of tidyverse to show the unique list in the column 
script - violin plot 
script - cowplot 

```{r data frames}
library(tidyverse)

#Import merged data from previous GSE108989 and GSE99254 analysis
crc<- read.csv("merged_all_CRC_GSE108989.csv", header=T) 
nsclc<- read.csv("merged_all_NSCLC_GSE99254.csv", header=T) 

#Sanity test
dim(crc)
dim(nsclc)

crc[c(1:5), c(1:12)]
nsclc[c(1:5), c(1:12)]

#Start from crc dataset
distinct(crc, Cluster)

#Note that Cluster in Tamatoa and published data is different

#In the paper
#CD4_C01-CCR7 Tn
#CD4_C10-FOXP3
#CD4_C11-IL10 Follicular Treg cells
#CD4_C12-CTLA4

#In Tamatoa (my current cluster info)
#CD4_C01-CCR7
#CD4_C10-IKZF2
#CD4_C11-IL10
#CD4_C12-CCR8

#Select the subset of CD4 cluster crom CRC data
crc.cd4<- crc %>% filter(Cluster=="CD4_C01-CCR7"|Cluster=="CD4_C10-IKZF2"|Cluster=="CD4_C11-IL10"|Cluster=="CD4_C12-CCR8")
distinct(crc.cd4, Cluster) #Now selected 4 clusters.
distinct(crc.cd4, SampleType)

#SampleType info from description from GSE108989 omnibus.
#PTC(CD8+ T cells from peripheral blood), NTC(CD8+T cells from adjacent normal colonrectal tissues) ,TTC (CD8+ T cells from tumor), 
#PTH(CD3+, CD4+ and CD25- T cells from peripheral blood),NTH(CD3+, CD4+ and CD25- T cells from adjacent normal colonrectal tissues),TTH(CD3+, CD4+ and CD25- T cells from tumor), 
#PTR(CD3+, CD4+ and CD25high T cells from peripheral blood), NTR(CD3+, CD4+ and CD25high Tcells from adjacent normal colonrectal tissues), TTR(CD3+, CD4+ and CD25high T cells from tumor), 
#PTY(CD3+, CD4+ and CD25mediate T cells from peripheral blood), NTY(CD3+, CD4+ and CD25mediate T cells from adjacent normal colonrectal tissues), TTY(CD3+, CD4+ and CD25medate T cells from tumor), 
#PP7(CD3+, CD4+ T cells from peripheral blood), NP7(CD3+, CD4+ T cells from adjacent normal colonrectal tissues), TP7(CD3+, CD4+ T cells from tumor).
#I combine all three normal, all three blood and all three tumor samples together into a individual categories. 

crc.p <- crc.cd4 %>%
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(SampleType=="PTR"|SampleType=="PTH"|SampleType=="PTY") #IL2RA is CD25 alias

crc.n <- crc.cd4 %>%
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(SampleType=="NTR"|SampleType=="NTH"|SampleType=="NTY")

crc.t <- crc.cd4 %>%
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(SampleType=="TTR"|SampleType=="TTH"|SampleType=="TTY") 



#Conduct the same analysis from NSCLC dataset 
distinct(nsclc, tissueType)

distinct(nsclc, majorCluster)
#In Paper
#CD4_C1-CCR7 Naive CD4
#CD4_C8-FOXP3 resting Treg
#CD4_C9-CTLA4 suppressive Treg

#In Tamatoa (Sampe populations were assigned. Good)
#CD4_C1-CCR7
#CD4_C8-FOXP3 
#CD4_C9-CTLA4

nsclc.cd4<- nsclc %>% filter(majorCluster=="CD4_C1-CCR7"|majorCluster=="CD4_C8-FOXP3"|majorCluster=="CD4_C9-CTLA4")
distinct(nsclc.cd4, majorCluster) #Now selected 3 clusters of interest.

distinct(nsclc.cd4, tissueType) #Show unique values (all clusters)

nsclc.p <- nsclc.cd4 %>%
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(tissueType == "Peripheral blood (P)") 

nsclc.n <- nsclc.cd4 %>%
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(tissueType == "Normal (N)") 

nsclc.t <- nsclc.cd4 %>%
  select(tissueType, majorCluster, CCR8, CXorf21,CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(tissueType == "Tumor-infiltrating (T)") 

```



```{r Pie chart}

library(ggplot2)
library(scales) #to format the labels in percent. Was not used here
library(ggrepel) #was not used here

#Pie Chart: cell number incidence of of n, p and t population of ALL cluster

df.crc <-data.frame(
  group = c("Peripheral", "Normal", "Tumor"),
  value = c(nrow(crc.p), nrow(crc.n), nrow(crc.t))
)

# Reordering plot based on Factor levels: https://sebastiansauer.github.io/ordering-bars/
# In ggplot2, defalt is alphabetical order when character. If factor, the "order of factor levels"" is used
df.crc$group <- factor(df.crc$group, levels=c("Peripheral", "Normal", "Tumor")) #Reorder plot

pie.crc <- ggplot(df.crc, aes(x="", y=value, fill=group))+
  geom_bar(width =1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("Tissue Origin of \n CRC CD4 T cells (all 4 clusters)") +
  xlab("# of Cells") + ylab("")+
  theme_bw()

plot(pie.crc)
#Save as 6x6


df.nsclc <-data.frame(
  group = c("Peripheral", "Normal", "Tumor"),
  value = c(nrow(nsclc.p), nrow(nsclc.n), nrow(nsclc.t))
)

df.nsclc$group <- factor(df.nsclc$group, levels=c("Peripheral", "Normal", "Tumor"))

pie.nsclc <- ggplot(df.nsclc, aes(x="", y=value, fill=group))+
  geom_bar(width =1, stat = "identity")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("Tissue Origin of \n NSCLC CD4T (all 3 clusters)") +
  xlab("# of Cells") + ylab("")+
  theme_bw()

plot(pie.nsclc)




```


```{r Analysis of individual clusters}
#Dissection of origin from individual clusters
#I have three layers of info. 1. Tissue, 2. Cluster, and 3. Value

## CRC dataset

#Calculate corresponding 12 values and assign to each

#Periphery
crc.p <- crc.cd4 %>%
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(SampleType=="PTR"|SampleType=="PTH"|SampleType=="PTY") 

distinct(crc.p, Cluster)

#Periphery&cluster
crc.p.C01.CCR7 <- crc.p %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C01-CCR7")

crc.p.C10.IKZF2 <- crc.p %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C10-IKZF2")

crc.p.C11.IL10 <- crc.p %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C11-IL10")

crc.p.C12.CCR8 <- crc.p %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C12-CCR8")



#Normal
crc.n <- crc.cd4 %>%
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(SampleType=="NTR"|SampleType=="NTH"|SampleType=="NTY")

#Normal&cluster
crc.n.C01.CCR7 <- crc.n %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C01-CCR7")

crc.n.C10.IKZF2 <- crc.n %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C10-IKZF2")

crc.n.C11.IL10 <- crc.n %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C11-IL10")

crc.n.C12.CCR8 <- crc.n %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C12-CCR8")


#Tumor
crc.t <- crc.cd4 %>%
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(SampleType=="TTR"|SampleType=="TTH"|SampleType=="TTY") 

#Tumor&cluster
crc.t.C01.CCR7 <- crc.t %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C01-CCR7")

crc.t.C10.IKZF2 <- crc.t %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C10-IKZF2")

crc.t.C11.IL10 <- crc.t %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C11-IL10")

crc.t.C12.CCR8 <- crc.t %>% 
  select(SampleType, Cluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(Cluster=="CD4_C12-CCR8")

#Generate a summary table
tissue <- rep(c("Peripheral", "Normal", "Tumor"), each = 4)
cluster <- rep(c("CD4_C01-CCR7","CD4_C10-IKZF2","CD4_C11-IL10","CD4_C12-CCR8"), 3)
value <- 
  c(nrow(crc.p.C01.CCR7), nrow(crc.p.C10.IKZF2),nrow(crc.p.C11.IL10), nrow(crc.p.C12.CCR8), nrow(crc.n.C01.CCR7), nrow(crc.n.C10.IKZF2), nrow(crc.n.C11.IL10), nrow(crc.n.C12.CCR8), nrow(crc.t.C01.CCR7), nrow(crc.t.C10.IKZF2), nrow(crc.t.C11.IL10), nrow(crc.t.C12.CCR8))

df.crc.all <- data.frame(tissue, cluster, value)
df.crc.all$tissue <- factor(df.crc.all$tissue, levels=c("Peripheral", "Normal", "Tumor")) #Reorder bar graph by factor-levels
factor(df.crc.all$tissue, levels=c("Peripheral", "Normal", "Tumor"))

library(hrbrthemes) #For theme, but did not use
library(viridis)
#https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html

bar.df.crc.all <- 
  ggplot(df.crc.all, aes(fill=cluster, y=value, x=tissue)) + 
  geom_bar(position="fill", stat="identity")+
  ggtitle("Tissue distribution of CD4 clusters in CRC")+
  xlab("")+
  ylab("")+
  scale_fill_viridis(discrete = T)+
  theme_bw()

plot(bar.df.crc.all)



## NSCLC dataset
#Calculate corresponding 9 values and assign to each, same as CRC

#Periphery
nsclc.p <- nsclc.cd4 %>%
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(tissueType == "Peripheral blood (P)")  

distinct(nsclc.p, majorCluster)

#Periphery&cluster
nsclc.p.C01.CCR7 <- nsclc.p %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C1-CCR7")

nsclc.p.C08.FOXP3 <- nsclc.p %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C8-FOXP3")

nsclc.p.C09.CTLA4 <- nsclc.p %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C9-CTLA4")



#Normal
nsclc.n <- nsclc.cd4 %>%
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(tissueType == "Normal (N)") 

#Normal&cluster
nsclc.n.C01.CCR7 <- nsclc.n %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C1-CCR7")

nsclc.n.C08.FOXP3 <- nsclc.n %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C8-FOXP3")

nsclc.n.C09.CTLA4 <- nsclc.n %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C9-CTLA4")


#Tumor
nsclc.t <- nsclc.cd4 %>%
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(tissueType == "Tumor-infiltrating (T)") 

#Tumor&cluster
nsclc.t.C01.CCR7 <- nsclc.t %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C1-CCR7")

nsclc.t.C08.FOXP3 <- nsclc.t %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C8-FOXP3")

nsclc.t.C09.CTLA4 <- nsclc.t %>% 
  select(tissueType, majorCluster, CCR8, CXorf21, CTLA4, FOXP3, IL2RA, TNFRSF9, TNFRSF8, EBI3, CCL22) %>%
  filter(majorCluster=="CD4_C9-CTLA4")


tissue <- rep(c("Peripheral", "Normal", "Tumor"), each = 3)
cluster <- rep(c("CD4_C1-CCR7","CD4_C8-FOXP3","CD4_C9-CTLA4"), 3)
value <- 
  c(nrow(nsclc.p.C01.CCR7), nrow(nsclc.p.C08.FOXP3),nrow(nsclc.p.C09.CTLA4), nrow(nsclc.n.C01.CCR7), nrow(nsclc.n.C08.FOXP3),nrow(nsclc.n.C09.CTLA4), nrow(nsclc.t.C01.CCR7), nrow(nsclc.t.C08.FOXP3),nrow(nsclc.t.C09.CTLA4))

df.nsclc.all <- data.frame(tissue, cluster, value)
df.nsclc.all$tissue <- factor(df.nsclc.all$tissue, levels=c("Peripheral", "Normal", "Tumor")) #Reorder bar graph by factor-levels


library(hrbrthemes) #For theme, bud did not use
library(viridis)
#https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html


bar.df.nsclc.all <- 
  ggplot(df.nsclc.all, aes(fill=cluster, y=value, x=tissue)) + 
  geom_bar(position="fill", stat="identity")+
  ggtitle("Tissue distribution of CD4 clusters in NSCLC")+
  xlab("")+
  ylab("")+
  scale_fill_viridis(discrete = T)+
  theme_bw()

plot(bar.df.nsclc.all)


```



```{r violet plot of individual gene: CCR8}

#Graph3 Violet plot of Normal, Peripheral and Tumor T cells (regardless of cluster)

#CCR8 distribution of CRC dataset.

crc.p$tissue <- rep("Peripheral", nrow(crc.p))
crc.n$tissue <- rep("Normal", nrow(crc.n)) #Need to add new tissue column to combine NTH etc. 
crc.t$tissue <- rep("Tumor", nrow(crc.t))
  
#now you can combine them, which maintain tissue infomration. 
df.crc.vplot <- rbind(crc.p, crc.n, crc.t)
df.crc.vplot$tissue <- factor(df.crc.vplot$tissue, levels=c("Peripheral", "Normal", "Tumor")) #Reorder bar graph by factor-levels


v.crc.ccr8 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=CCR8, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCR8 exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

plot(v.crc.ccr8)

#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.ccr8.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=CCR8, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("CCR8 exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important

plot(v.crc.ccr8.bycluster)




#NSCLC
nsclc.p$tissue <- rep("Peripheral", nrow(nsclc.p))
nsclc.n$tissue <- rep("Normal", nrow(nsclc.n)) #Assignment is redundant but did it anyway to have same labeling. 
nsclc.t$tissue <- rep("Tumor", nrow(nsclc.t))

df.nsclc.vplot <- rbind(nsclc.p, nsclc.n, nsclc.t)
df.nsclc.vplot$tissue <- factor(df.nsclc.vplot$tissue, levels=c("Peripheral", "Normal", "Tumor")) #Reorder bar graph by factor-levels

v.nsclc.ccr8 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=CCR8, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCR8 exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

plot(v.nsclc.ccr8)

v.nsclc.ccr8.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=CCR8, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("CCR8 exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important


plot(v.nsclc.ccr8.bycluster)

```


```{r violet plot of individual gene: CTLA4, FOXP3, IL2RA and others}

#These plot depends on df.crc.vplot table generated in previous section

#CTLA4

v.crc.ctla4 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=CTLA4, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CTLA4 exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

plot(v.crc.ctla4)

#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.ctla4.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=CTLA4, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("CTLA4 exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important

plot(v.crc.ctla4.bycluster)


v.nsclc.ctla4 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=CTLA4, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CTLA4 exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

plot(v.nsclc.ctla4)

v.nsclc.ctla4.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=CTLA4, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("CTLA4 exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important

plot(v.nsclc.ctla4.bycluster)







#FOXP3

v.crc.foxp3 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=FOXP3, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("FOXP3 exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

plot(v.crc.foxp3)

#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.foxp3.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=FOXP3, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("FOXP3 exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important

plot(v.crc.foxp3.bycluster)


v.nsclc.foxp3 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=FOXP3, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("FOXP3 exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

plot(v.nsclc.foxp3)

v.nsclc.foxp3.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=FOXP3, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("FOXP3 exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important

plot(v.nsclc.foxp3.bycluster)




#IL2RA (CD25)

v.crc.il2ra <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=IL2RA, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("IL2RA(CD25) exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.il2ra.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=IL2RA, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("IL2RA(CD25) exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important



v.nsclc.il2ra <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=IL2RA, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("IL2RA(CD25) exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()

v.nsclc.il2ra.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=IL2RA, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("IL2RA(CD25) exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important







#TNFRSF9 (4-1BB)

v.crc.tnfrsf9 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=TNFRSF9, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("TNFRSF9(4-1BB) exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.tnfrsf9.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=TNFRSF9, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("TNFRSF9(4-1BB) exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important


v.nsclc.tnfrsf9 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=TNFRSF9, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("TNFRSF9(4-1BB) exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


v.nsclc.tnfrsf9.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=TNFRSF9, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("TNFRSF9(4-1BB) exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important




# TNFRSF8

v.crc.tnfrsf8 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=TNFRSF8, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("TNFRSF8 exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.tnfrsf8.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=TNFRSF8, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("TNFRSF8 exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important


v.nsclc.tnfrsf8 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=TNFRSF8, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("TNFRS8 exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


v.nsclc.tnfrsf8.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=TNFRSF8, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("TNFRSF8 exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important




#EBI3

v.crc.ebi3 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=EBI3, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("EBI3 exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.ebi3.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=EBI3, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("EBI3 exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important


v.nsclc.ebi3 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=EBI3, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("EBI3 exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


v.nsclc.ebi3.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=EBI3, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("EBI3 exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important



#CCL22


v.crc.ccl22 <- 
  ggplot(df.crc.vplot, aes(x=tissue, y=CCL22, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCL22 exp distribution in CRC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


#This plot doesnt need to re-order. Alphabetical order is good enough
v.crc.ccl22.bycluster <- 
ggplot(crc.cd4, aes(x=Cluster, y=CCL22, fill=Cluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("CCL22 exp distribution in CRC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important


v.nsclc.ccl22 <- 
  ggplot(df.nsclc.vplot, aes(x=tissue, y=CCL22, fill=tissue)) + 
  geom_violin(trim=FALSE)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCL22 exp distribution in NSCLC by tissue")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()


v.nsclc.ccl22.bycluster <- 
ggplot(nsclc.cd4, aes(x=majorCluster, y=CCL22, fill=majorCluster))+
  geom_violin(trim=FALSE)+
  scale_fill_viridis(discrete = T)+
  ggtitle("CCL22 exp distribution in NSCLC by cluster")+
  xlab("") + ylab("Log2(Norm.TPM)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1)) #Note that theme_bw will override theme. Order here is important


```


```{r, echo=FALSE,results='hide',fig.keep='all'}

#Final Plots summary

plot(pie.crc)
plot(pie.nsclc)
plot(bar.df.crc.all)
plot(bar.df.nsclc.all)
plot(v.crc.ccr8)
plot(v.crc.ccr8.bycluster)
plot(v.nsclc.ccr8)
plot(v.nsclc.ccr8.bycluster)



#Treg_genes

#CTLA4
plot(v.crc.ctla4)
plot(v.crc.ctla4.bycluster)
plot(v.nsclc.ctla4)
plot(v.nsclc.ctla4.bycluster)

#FOXP3
plot(v.crc.foxp3)
plot(v.crc.foxp3.bycluster)
plot(v.nsclc.foxp3)
plot(v.nsclc.foxp3.bycluster)

#IL2RA (CD25)
plot(v.crc.il2ra)
plot(v.crc.il2ra.bycluster)
plot(v.nsclc.il2ra)
plot(v.nsclc.il2ra.bycluster)


#Volcano_Hits

#TNFRSF9 (4-1BB)
plot(v.crc.tnfrsf9)
plot(v.crc.tnfrsf9.bycluster)
plot(v.nsclc.tnfrsf9)
plot(v.nsclc.tnfrsf9.bycluster)


#TNFRSF8
plot(v.crc.tnfrsf8)
plot(v.crc.tnfrsf8.bycluster)
plot(v.nsclc.tnfrsf8)
plot(v.nsclc.tnfrsf8.bycluster)

#EBI3
plot(v.crc.ebi3)
plot(v.crc.ebi3.bycluster)
plot(v.nsclc.ebi3)
plot(v.nsclc.ebi3.bycluster)


#CCL22
plot(v.crc.ccl22)
plot(v.crc.ccl22.bycluster)
plot(v.nsclc.ccl22)
plot(v.nsclc.ccl22.bycluster)

```




## cowplot to generate final figure

```{r cowplot}
#cowplot to show all results side-by-side
#https://stackoverflow.com/questions/1249548/side-by-side-plots-with-ggplot2

library(cowplot)

#Generate a combined plot
p1<-
  plot_grid(pie.crc, pie.nsclc, bar.df.crc.all, bar.df.nsclc.all, v.crc.ccr8, v.nsclc.ccr8, v.crc.ccr8.bycluster, v.nsclc.ccr8.bycluster, ncol=2, labels = "AUTO")

plot(p1)

ggsave("GSE99254&GSE108989_CCR8_plots_combined.pdf", p1, width = 12, height = 14)




p2<-
  plot_grid(v.crc.ctla4, v.nsclc.ctla4, v.crc.ctla4.bycluster, v.nsclc.ctla4.bycluster, v.crc.foxp3, v.nsclc.foxp3, v.crc.foxp3.bycluster, v.nsclc.foxp3.bycluster, v.crc.il2ra, v.nsclc.il2ra, v.crc.il2ra.bycluster, v.nsclc.il2ra.bycluster,ncol=2, labels = "AUTO")

plot(p2)

ggsave("GSE99254&GSE108989_Treg_Genes_of_Interests.pdf", p2, width = 12, height = 21)



p3<-
  plot_grid(v.crc.tnfrsf9, v.nsclc.tnfrsf9, v.crc.tnfrsf9.bycluster, v.nsclc.tnfrsf9.bycluster, v.crc.tnfrsf8, v.nsclc.tnfrsf8, v.crc.tnfrsf8.bycluster, v.nsclc.tnfrsf8.bycluster, v.crc.ebi3, v.nsclc.ebi3, v.crc.ebi3.bycluster, v.nsclc.ebi3.bycluster, v.crc.ccl22, v.nsclc.ccl22, v.crc.ccl22.bycluster, v.nsclc.ccl22.bycluster, ncol=2, labels = "AUTO")

plot(p3)

ggsave("GSE99254&GSE108989_Other_Genes_of_Interests.pdf", p3, width = 12, height = 28)


```





```{r Tutorial, results="hide"}

#Tutorial note.
# Filter multiple using %in%: https://stackoverflow.com/questions/25647470/filter-multiple-values-on-a-string-column-in-dplyr

categories <- c("PTR", "NTR", "TTR")
crc.plot <- crc %>%
  select(SampleType, CCR8) %>%
  filter(SampleType %in% categories) #1042 obs to 889 obs

head(crc.plot)

v.crc <- ggplot(crc.plot, aes(x=SampleType, y=CCR8, color=SampleType)) + 
  geom_violin(trim=FALSE)+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  ggtitle("CCR8 expression in CRC")+
  theme_bw()




```


